Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations

### 2021-11-11 analysis/paper/paper.Rmd

```{r cmr-effort-data}
cmr_df = read_csv("../data/raw_data/deadwater_cmr_effort_20210614.csv") %>%
  clean_names()

@@

>>>>

```{r cmr-effort-data}
# cmr_df = read_csv(here("analysis/data/raw_data/deadwater_cmr_effort_20210614.csv")) %>%
cmr_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
  clean_names() %>%
  filter(count > 0) %>%
  # pull out only Northern Pikeminnow
  filter(species == "Northern Pikeminnow") %>%
  # mutate(across(date,
  #               mdy)) %>%
  mutate(across(date,
                ymd)) %>%
  mutate(across(c(start_time, end_time),
                ~ ymd_hms(paste(year(date), month(date), day(date), hour(.), minute(.), second(.))))) %>%
  # split into sampling periods
  mutate(event_name = if_else(month(date) > 8,
                              paste("Fall", year(date), sep = "_"),
                              paste("Spring", year(date), sep = "_")))

effort_df = cmr_df %>%
  select(event_name,
         site,
         date,
         crew,
         method,
         equipment,
         start_time:total_effort,
         total_day_effort) %>%
  distinct() %>%
  group_by(event_name,
           date) %>%
  summarise(across(total_effort,
                   sum),
            .groups = "drop")

cpue_df = effort_df %>%
  # add how many pikeminnow were caught
  left_join(cmr_df %>%
              filter(capture_type != "Recapture") %>%
              group_by(event_name, species, date) %>%
              summarise(n_fish = n(),
                        .groups = "drop")) %>%
  mutate(cpue = n_fish / total_effort)

@@

```{r lincoln-petersen-estimator}
# using Lincoln-Petersen model, combining first 4 and last four days
# define mark and recapture periods
cap_periods = tibble(event_name = "Fall_2019",
                     cap_period = "Mark",
                     int = interval(start = ymd('20191112'),
                                    end = ymd('20191114'))) %>%
  bind_rows(tibble(event_name = "Fall_2019",
                   cap_period = "Recapture",
                   int = interval(start = ymd('20191119'),
                                  end = ymd('20191121')))) %>%
  bind_rows(tibble(event_name = "Fall_2020",
                   cap_period = "Mark",
                   int = interval(start = ymd('20201020'),
                                  end = ymd('20201023')))) %>%
  bind_rows(tibble(event_name = "Fall_2020",
                   cap_period = "Recapture",
                   int = interval(start = ymd('20201027'),
                                  end = ymd('20201030'))))

cmr_df$cap_period = NA_character_
for(i in 1:nrow(cap_periods)) {
  my_int = cap_periods$int[i]
  index = which(cmr_df$date %within% my_int)
  cmr_df$cap_period[index] = cap_periods$cap_period[i]
}

LP_mod = cmr_df %>%
  filter(species == "Northern Pikeminnow") %>%
  # filter out spring 2021 (no recapture)
  filter(event_name != "Spring_2021") %>%
  mutate(mr_type = if_else(cap_period == "Mark" & fish_status == "Live",
                           "M",
                           if_else(cap_period == "Recapture",
                                   if_else(capture_type == "Mark",
                                           "C",
                                           if_else(capture_type == "Recapture",
                                                   "R",
                                                   NA_character_)),
                                   NA_character_))) %>%
  filter(!is.na(mr_type)) %>%
  group_by(event_name, species) %>%
  summarise(M = sum(mr_type == "M"),
            n = sum(mr_type %in% c("C", "R")),
            m = sum(mr_type == 'R'),
            .groups = "drop") %>%
  nest(data = -c(event_name, species)) %>%
  crossing(model = c("Petersen",
                     "Chapman")) %>%
  mutate(mr_mod = map2(data,
                       model,
                      .f = function(x, y) {
                        with(x,
                             mrClosed(M = M,
                                      n = n,
                                      m = m,
                                      method = y))
                      }),
         N_summ = map(mr_mod,
                      .f = summary,
                      incl.SE = T),
         CI = map(mr_mod,
                  .f = confint)) %>%
  mutate(N = map_dbl(N_summ,
                     .f = function(x) x[1]),
         SE = map_dbl(N_summ,
                      .f = function(x) x[2]),
         N.lci = map_dbl(CI,
                         .f = function(x) x[1]),
         N.uci = map_dbl(CI,
                         .f = function(x) x[2])) %>%
  select(-N_summ, -CI)
```

```{r schnabel-estimator}
# using Schnabel model
N_mods = cmr_df %>%
  filter(species == "Northern Pikeminnow") %>%
  # filter out spring 2021 (no recapture)
  filter(event_name != "Spring_2021") %>%
  group_by(event_name, species, date, fish_status, capture_type) %>%
  summarise(n_fish = n(),
            .groups = "drop") %>%
  group_by(event_name, species, date) %>%
  summarise(n = sum(n_fish),
            m = sum(n_fish[capture_type == "Recapture"]),
            R = sum(n_fish[fish_status == "Live"]),
            dead_fish = sum(n_fish[fish_status == "Dead"]),
            .groups = "drop") %>%
  group_by(event_name, species) %>%
  nest() %>%
  mutate(mod = map(data,
                   .f = function(x) {
                     with(x,
                          mrClosed(n = n,
                                   m = m,
                                   R = R,
                                   method = "Schnabel",
                                   chapman.mod = TRUE))
                   }),
         N = map_dbl(mod,
                     .f = summary),
         CI = map(mod,
                  .f = confint))

N_mods %>%
  select(species, N, CI) %>%
  mutate(N.lci = map_dbl(CI,
                         .f = function(x) x[1]),
         N.uci = map_dbl(CI,
                         .f = function(x) x[2])) %>%
  select(-CI)
```
